Loop Invariant Code Motion in PyPy's Tracing JIT

### 2011-06-16 talk/iwtc11/paper.tex
oops

\section{Trace Optimizations}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=1]{figures/overview}
\end{center}
\end{figure}

XXX find reference

Loop peeling is achieved by appending a copy of the traced iteration at
the end of the loop. The copy  is inlined to make the two parts form a
consistent two iteration trace. 
The first part (called preamble) finishes with the jump the the second part
(called peeled loop). The second part ends up with the jump to itself. This way
the preamble will be executed only once while the peeled loop will
be used for every other iteration.
The trace from Figure~\ref{fig:unopt-trace} would after this operation become
the trace in Figure~\ref{fig:peeled-trace}. Line 1-13 shows the
preamble while line 15-27 shows the peeled loop.

@@

Each operation in the trace is inlined in order.
To inline an operation $v=\text{op}\left(A_1, A_2, \cdots, A_{|A|}\right)$
a new variable, $\hat v$ is introduced. The inlined operation will
produce $\hat v$ using
\begin{equation}
  \hat v = \text{op}\left(m\left(A_1\right), m\left(A_2\right), 
    \cdots, m\left(A_{|A|}\right)\right) . 
\end{equation}
Before the
next operation is inlined, $m$ is extend by assigning $m\left(v\right) = \hat
v$. For the example above, after all the operations have been inlined we have
\begin{equation}
  %\left\{
    \begin{array}{lcl}
      m\left(p_0\right) &=& p_0 \\
      m\left(p_1\right) &=& p_5 \\
      m\left(i_2\right) &=& i_6 \\
      m\left(i_3\right) &=& i_7 \\
      m\left(i_4\right) &=& i_8 \\
      m\left(p_5\right) &=& p_9 \\
    \end{array}
  %\right.
  .
\end{equation}

@@

XXX should we have a mention where in the previous papers those optimizations
are discussed? Is the previous XXX precisely about this?

>>>>

\section{Making Trace Optimizations Loop Aware}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=1]{figures/overview}
\end{center}
\caption{Overview of Loop Peeling}
\label{fig:overview}
\end{figure}

XXX find reference of prior work on this

Loop peeling is achieved by appending a copy of the traced iteration at
the end of the loop. See Figure~\ref{fig:overview}
The first part (called \emph{preamble}) finishes with the jump the the second part
(called the \emph{peeled loop}). The second part end with the jump to itself. This way
the preamble will be executed only once while the peeled loop will
be used for every further iteration.

@@

Each operation in the trace is copied in order.
To copy an operation $v=\text{op}\left(A_1, A_2, \cdots, A_{|A|}\right)$
a new variable, $\hat v$ is introduced. The copied operation will
return $\hat v$ using
\begin{equation}
  \hat v = \text{op}\left(m\left(A_1\right), m\left(A_2\right), 
    \cdots, m\left(A_{|A|}\right)\right) . 
\end{equation}
Before the
next operation is copied, $m$ is extend by assigning $m\left(v\right) = \hat
v$. For the example above, after all the operations have been copied we have
\begin{equation}
  %\left\{
    \begin{array}{lcl}
      m\left(p_0\right) &=& p_0 \\
      m\left(p_1\right) &=& p_5 \\
      m\left(i_2\right) &=& i_6 \\
      m\left(i_3\right) &=& i_7 \\
      m\left(i_4\right) &=& i_8 \\
      m\left(p_5\right) &=& p_9 \\
    \end{array}
  %\right.
  .
\end{equation}

The trace from Figure~\ref{fig:unopt-trace} would after this operation become
the trace in Figure~\ref{fig:peeled-trace}. Line 1-13 shows the
preamble while line 15-27 shows the peeled loop.

\section{Interaction of Optimizations with Loop Peeling}

@@

XXX should we have a mention where in the previous papers those optimizations
are discussed?
